---
title: "Titanic_Arbeit"
author: "Dona Olivier D. Whannou"
date: "2026-03-30"
output: pdf_document
---

```{r setup, include=FALSE}

library(ggplot2)
library(tidyverse)
library(caret)
library(corrplot)
library(MASS)

```
#INTRODUCTION
The sinking of the Titanic is one of the most infamous shipwrecks in history.
In this challenge, we ask you to build a predictive model that answers the question: “what sorts of people were more likely to survive?” using passenger data (ie name, age, gender, socio-economic class, etc).


#DATA UNDERSTANDING AND VISUALIZATION


```{r }
T<-read.csv("train.csv")
T2 <- read.csv("test.csv")
str(T)
str(T2)
summary(T)

```

```{r }


ggplot(T, aes(x = factor(Pclass), fill = factor(Survived))) +
  geom_bar(position = "fill") +
  labs(
    title = "Survival by Class",
    x = "Passenger Class",
    y = "Count",
    fill = "Survived"
  ) +
  scale_fill_manual(values = c("brown1", "deepskyblue"),
                    labels = c("Died", "Survived")) +
  theme_minimal()
#People in third class had a significantly lower survival probability than those in other classes. Around 25% passenger in class 3 survived in compare with 62,5% in class1.

```


```{r }
ggplot(T, aes(x = Sex, fill = factor(Survived))) +
  geom_bar(position = "fill") +
  labs(
    title = "Survival Rate by Sex",
    x = "Sex",
    y = "Proportion",
    fill = "Survived"
  ) +
  scale_fill_manual(values = c("brown1", "deepskyblue"),
                    labels = c("Died", "Survived")) +
  theme_minimal()
```


```{r }
T$AgeGroup <- cut(
  T$Age,
  breaks = seq(0, 80, by = 5),
  include.lowest = TRUE
)
table(T$AgeGroup)

ggplot(T, aes(x = AgeGroup, fill = factor(Survived))) +
  geom_bar(position = "fill") +
  labs(
    title = "Survival Rate by AgeGroup",
    x = "AgeGroup",
    y = "Proportion",
    fill = "Survived"
  ) +
  scale_fill_manual(values = c("brown1", "deepskyblue"),
                    labels = c("Died", "Survived")) +
  theme_minimal()
```
# MISSING VALUES, AMPUTATION AND FEATURE ENGINEERING 


```{r }
colSums(is.na(T))
round(colSums(is.na(T)) / nrow(T) * 100, 2)
#only Age had 177 missing values. Since 17% of the values are missing,I can replace them with the mean or the median.

medianage<-median(T$Age,na.rm=TRUE)
meanmedian<-mean(T$Age, na.rm = TRUE)
T$Age[is.na(T$Age)]<-medianage

#Cabin 
sum(T$Cabin=="") #687 of 891
T$CabinKnown <- ifelse(T$Cabin == "", 0, 1)

#In the cabin section, there are a lot of missing values right now. 
#I don't know why #they haven't been entered (I'll look into it later).
#For now, I've assigned 0 to the missing values and 1 to the opposite.
#I'll create a model that distinguishes between cabins and non-cabins.
#Because I think the rate of missing values is too high.




#### Test values
colSums(is.na(T2))
round(colSums(is.na(T2)) / nrow(T2) * 100, 2)
# only Age had 177 missing values.
# Since 17% of the values are missing, I can replace them with the mean or the median.

medianage<-median(T2$Age,na.rm=TRUE)
meanmedian<-mean(T2$Age, na.rm = TRUE)
T2$Age[is.na(T2$Age)]<-medianage

#Cabin 
sum(T2$Cabin=="") #687 of 891
T2$CabinKnown <- ifelse(T2$Cabin == "", 0, 1)
```

```{r }
#The encoding of sex and Embarkation to see the correlation wih another varibles (likesurvived, pclass...)
#sex male(0), female(1)
T$Sexnew<- ifelse(T$Sex=="male",0,1)
# Port of Embarkation	C = Cherbourg(1), Q = Queenstown(2), S = Southampton(0)
T$Embarkednew<- ifelse(T$Embarked=="S",0,
                       ifelse(T$Embarked=="C",1,2))


#####
T2$Sexnew<- ifelse(T2$Sex=="male",0,1)
# Port of Embarkation	C = Cherbourg(1), Q = Queenstown(2), S = Southampton(0)

T2$Embarkednew<- ifelse(T2$Embarked=="S",0,
                       ifelse(T2$Embarked=="C",1,2))


#######
numeric_vars <- T[, sapply(T, is.numeric)]
cor_matrix <- round(cor(numeric_vars, use = "complete.obs"),3)
cor_matrix
corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.8)
```

```{r }
#work on the another character
T$title<- sub(".*, (.*?)\\..*", "\\1",T$Name)
round(prop.table(table(T$title)),3)


#the important title are Mr, Miss(Miss,Mlle,Mrs,mme), Master and rare(capt,col,Don,Dr,Jonkheer, Major,sir, rev,teh countess)
T$title_group <- ifelse(T$title %in% c("Mr"), "Mr",
                  ifelse(T$title %in% c("Miss","Mlle","Ms"), "Miss",
                  ifelse(T$title %in% c("Mrs","Mme"), "Mrs",
                  ifelse(T$title == "Master", "Master", "Rare"))))


#test values
T2$title<- sub(".*, (.*?)\\..*", "\\1",T2$Name)
round(prop.table(table(T2$title)),3)

#the important title are Mr, Miss(Miss,Mlle,Mrs,mme), Master and rare(capt,col,Don,Dr,Jonkheer, Major,sir, rev,teh countess)

T2$title_group <- ifelse(T2$title %in% c("Mr"), "Mr",
                  ifelse(T2$title %in% c("Miss","Mlle","Ms"), "Miss",
                  ifelse(T2$title %in% c("Mrs","Mme"), "Mrs",
                  ifelse(T2$title == "Master", "Master", "Rare"))))
```


###  Logistic regression

```{r }
model <- glm(Survived ~ Pclass + Sexnew + Age + Fare+Embarkednew +CabinKnown,
             data = T,
             family = binomial)

pred <- predict(model, newdata = T, type = "response")
pred_class <- ifelse(pred > 0.57, 1, 0) #1= survived 0=died

train_eval<- table(Predicted = pred_class, Actual = T$Survived)
# train evaluation

TP1 <- train_eval["1","1"]
TN1 <- train_eval["0","0"]
FP1 <- train_eval["1","0"]
FN1 <- train_eval["0","1"]
Accuracy <-(TP1+TN1)/(TP1+FN1+TN1+FP1)
ConfusionMatrix<- confusionMatrix(as.factor(pred_class), as.factor(T$Survived))
print(ConfusionMatrix)
```




```{r }
#Using the data from the Titanic passenger manifest (test data),
#I have compiled a list of people who might have survived based on 
#my logistic regression model trained above.
#pred <- predict(model, newdata = T2, type = "response")
#pred_class <- ifelse(pred > 0.57, 1, 0)

#T2$survived <- pred_class
#submission <- T2[, c("PassengerId", "Survived")]


```



